---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  out.width = 600
)
```

# deltatest: Statistical Hypothesis Testing Using the Delta Method for Online A/B Testing

<!-- badges: start -->
[![R-CMD-check](https://github.com/hoxo-m/deltatest/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/hoxo-m/deltatest/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

## 1. Overview

Online A/B testing poses a significant practical challenge: the randomization unit often differs from the analysis unit. For instance, while control and treatment groups are typically assigned at the user level, evaluation metrics are frequently measured at a more detailed level, such as the click-through rate per page view. In this scenario, the randomization unit is the user, but the analysis unit is the page view. 

This discrepancy raises concerns when performing statistical hypothesis tests, as within-user correlation may be present. Specifically, a single user can generate multiple page views, and some users may inherently have a higher likelihood of clicking than others, which may potentially violate the i.i.d. assumption required for hypothesis tests.

To address this issue, Deng et al. (2018) proposes a statistical hypothesis testing method using the Delta method. This package has been developed to easily execute that method.

First, we prepare a data frame that includes the number of clicks and page views aggregated for each user. This data frame also contains a column indicating whether each user was assigned to the control or treatment group.

```{r prepare_data}
library(dplyr)

n_user <- 2000

set.seed(314)
data <- deltatest::generate_dummy_data(n_user) |> 
  mutate(group = if_else(group == 0, "control", "treatment")) |>
  group_by(user_id, group) |> 
  summarise(clicks = sum(metric), pageviews = n(), .groups = "drop")
data
```

To perform a statistical hypothesis test using the Delta method on this data, as follows:

```{r}
library(deltatest)

deltatest(data, clicks / pageviews, by = group)
```

## 2. Installation

You can install the development version of deltatest from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("hoxo-m/deltatest")
```

## Example

```{r example}
library(dplyr)
library(deltatest)

n_user <- 2000

set.seed(314)
df <- generate_dummy_data(n_user) |> 
  mutate(group = if_else(group == 0L, "control", "treatment")) |>
  group_by(user_id, group) |> 
  summarise(click = sum(metric), pageview = n(), .groups = "drop")
df

deltatest(df, click / pageview, by = group)
```

```{r cache=TRUE}
library(ggplot2)

set.seed(314)

p_values <- NULL
for (i in 1:5000) {
  df <- generate_dummy_data(n_user) |> 
    group_by(group) |> 
    summarise(click = sum(metric), pageview = n(), .groups = "drop")
  
  result <- prop.test(df$click, df$pageview, correct = FALSE)
  
  p_values[i] <- result$p.value
}

df <- data.frame(p_value = p_values) |>
  mutate(range = cut(p_value, breaks = seq(0, 1, by = 0.05))) |>
  group_by(range) |>
  summarise(p = factor(ceiling(max(p_value) * 20) / 20), n = n()) |>
  mutate(prop = n / sum(n))
ggplot(df, aes(p, prop)) +
  geom_col() +
  geom_hline(yintercept = 0.05, color = "red") +
  scale_y_continuous(breaks = seq(0, 1, by = 0.02), minor_breaks = NULL) +
  xlab("p-value") + ylab("proportion")
```

```{r cache=TRUE}
set.seed(314)

p_values <- NULL
for (i in 1:5000) {
  df <- generate_dummy_data(n_user) |> 
    group_by(user_id, group) |> 
    summarise(click = sum(metric), pageview = n(), .groups = "drop")
  
  result <- deltatest(df, click / pageview, by = group)
  
  p_values[i] <- result$p.value
}

df <- data.frame(p_value = p_values) |>
  mutate(range = cut(p_value, breaks = seq(0, 1, by = 0.05))) |>
  group_by(range) |>
  summarise(p = factor(ceiling(max(p_value) * 20) / 20), n = n()) |>
  mutate(prop = n / sum(n))
ggplot(df, aes(p, prop)) +
  geom_col() +
  geom_hline(yintercept = 0.05, color = "red") +
  scale_y_continuous(breaks = seq(0, 1, by = 0.01), minor_breaks = NULL) +
  xlab("p-value") + ylab("proportion")
```

## Related Work

- [tidydelta: Estimation of Standard Errors using Delta Method](https://cran.r-project.org/package=tidydelta)

## References

- Deng, A., Knoblich, U., & Lu, J. (2018). Applying the Delta Method in Metric
Analytics: A Practical Guide with Novel Ideas. *Proceedings of the 24th ACM
SIGKDD International Conference on Knowledge Discovery & Data Mining.*
[doi:10.1145/3219819.3219919](https://doi.org/10.1145/3219819.3219919)
- Deng, A., Lu, J., & Litz, J. (2017). Trustworthy Analysis of Online A/B Tests:
Pitfalls, challenges and solutions. *Proceedings of the Tenth ACM International
Conference on Web Search and Data Mining.*
[doi:10.1145/3018661.3018677](https://doi.org/10.1145/3018661.3018677)
- id:sz_dr (2018). Calculating the mean and variance of the ratio of random
variables using the Delta method [in Japanese]. *If you are human, think more
now.* https://www.szdrblog.info/entry/2018/11/18/154952
