---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = 400,
  message = FALSE
)
```

# deltatest: Statistical Hypothesis Testing Using the Delta Method for Online A/B Testing

<!-- badges: start -->
[![R-CMD-check](https://github.com/hoxo-m/deltatest/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/hoxo-m/deltatest/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

## 1. Overview

In online A/B testing, we often face a significant practical challenge: the randomization unit differs from the analysis unit. Typically, control and treatment groups are assigned at the user level, while metrics—such as click-through rate—are measured at a finer level (e.g., per page view). In this scenario, the randomization unit is a user, while the analysis unit is a page view.

This discrepancy raises concerns for statistical hypothesis testing, which assumes that data points are independent and identically distributed (i.i.d.). Specifically, a single user can generate multiple page views, and each user may have a different probability of clicking. As a result, the data may exhibit within-user correlation, thereby violating the i.i.d. assumption.

When the standard Z-test is applied to such correlated data, the resulting  p-values do not follow the expected uniform distribution under the null hypothesis. As a result, smaller p-values tend to occur more frequently despite there being no true difference, increasing the risk of falsely detecting a significant difference.

```{r p-values-from-z-test, echo=FALSE, fig.height=3, fig.width=4}
library(dplyr)
library(ggplot2)

file <- "data-raw/p_values_from_standard_Z_test.rds"

p_values <- 
  if (file.exists(file)) {
    readRDS(file)
  } else {
    source("data-raw/compute_p_values.R")
    readRDS(file)
  }

df <- data.frame(p_value = p_values) |>
  mutate(range = cut(p_value, breaks = seq(0, 1, by = 0.05))) |>
  group_by(range) |>
  summarise(p = factor(ceiling(max(p_value) * 20) / 20), n = n()) |>
  mutate(prop = n / sum(n))

ggplot(df, aes(p, prop)) +
  geom_col() +
  geom_hline(yintercept = 0.05, color = "red") +
  scale_y_continuous(breaks = seq(0, 1, by = 0.05)) +
  xlab("p-value") + ylab("proportion") +
  # ggtitle("Standard Z-test for data with within-user correlation") +
  ggtitle("p-values from standard Z-test on correlated data") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1), 
        plot.title = element_text(size = 11))
```

To address this issue, Deng et al. (2018) proposed a statistical hypothesis testing method that replaces the standard variance estimation formula in the Z-test with an approximate formula using the Delta method, which can account for within-user correlation. This package has been developed to make it easy to execute that method.

First, we prepare a data frame that includes the number of clicks and page views aggregated for each user. This data frame also contains a column indicating whether each user was assigned to the control or treatment group.

```{r prepare_data}
library(dplyr)

n_user <- 2000

set.seed(314)
data <- deltatest::generate_dummy_data(n_user) |> 
  mutate(group = if_else(group == 0, "control", "treatment")) |>
  group_by(user_id, group) |> 
  summarise(clicks = sum(metric), pageviews = n(), .groups = "drop")

data
```

To perform a statistical hypothesis test using the Delta method on this data, as follows:

```{r execute}
library(deltatest)

deltatest(data, clicks / pageviews, by = group)
```

The Z-test using the Delta method yields p-values that follow the expected uniform distribution under the null hypothesis, even in the presence of correlation within the data.

```{r p-values-from-delta-method, echo=FALSE, fig.height=3, fig.width=4}
p_values <- readRDS("data-raw/p_values_from_Delta_meethod.rds")

df <- data.frame(p_value = p_values) |>
  mutate(range = cut(p_value, breaks = seq(0, 1, by = 0.05))) |>
  group_by(range) |>
  summarise(p = factor(ceiling(max(p_value) * 20) / 20), n = n()) |>
  mutate(prop = n / sum(n))

ggplot(df, aes(p, prop)) +
  geom_col() +
  geom_hline(yintercept = 0.05, color = "red") +
  scale_y_continuous(breaks = seq(0, 1, by = 0.01)) +
  xlab("p-value") + ylab("proportion") +
  ggtitle("p-values from Z-test with Delta method on correlated data") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1), 
        plot.title = element_text(size = 10))
```

## 2. Installation

You can install the development version of deltatest from [GitHub](https://github.com/) with:

```{r install_github, eval=FALSE}
# install.packages("remotes")
remotes::install_github("hoxo-m/deltatest")
```

## 3. Details

### 3.1 Basic Usage

The **deltatest** package provides the `deltatest` function to perform the statistical hypothesis testing using the Delta method, as proposed by Deng et al. (2018).

To run `deltatest`, you need to prepare a data frame.
It must include columns for the numerator and denominator of your metric, aggregated by each randomization unit (typically, the randomization unit is a user).
For example:

- If the metric is CTR, the numerator is the number of clicks, and the denominator is the number of page views.
- If the metric is CVR, the numerator is the number of converted sessions, and the denominator is the number of sessions.

The `generate_dummy_data` function provides 

```{r generate_dummy_data}
library(dplyr)

n_user <- 2000

data <- deltatest::generate_dummy_data(n_user) |>
  group_by(user_id, group) |>
  summarise(clicks = sum(metric), pageviews = n(), .groups = "drop")

data
```


### 3.2 Expression Styles

1. Expression (NSE; non-standard evaluation)

```{r, eval=FALSE}
deltatest(data, clicks / pageviews, by = group)
```

2. Standard formula

```{r, eval=FALSE}
deltatest(data, clicks / pageviews ~ group)
```

3. Lambda formula

```{r, eval=FALSE}
deltatest(data, ~ clicks / pageviews, by = group)
```

- With calculation (for all styles)

```{r, eval=FALSE}
deltatest(data, pos / (pos + neg), by = group)
```

### 3.3 Information

## 4. Related Work

- [tidydelta: Estimation of Standard Errors using Delta Method](https://cran.r-project.org/package=tidydelta)

## 5. References

- Deng, A., Knoblich, U., & Lu, J. (2018). Applying the Delta Method in Metric
Analytics: A Practical Guide with Novel Ideas. *Proceedings of the 24th ACM
SIGKDD International Conference on Knowledge Discovery & Data Mining.*
[doi:10.1145/3219819.3219919](https://doi.org/10.1145/3219819.3219919)
- Deng, A., Lu, J., & Litz, J. (2017). Trustworthy Analysis of Online A/B Tests:
Pitfalls, challenges and solutions. *Proceedings of the Tenth ACM International
Conference on Web Search and Data Mining.*
[doi:10.1145/3018661.3018677](https://doi.org/10.1145/3018661.3018677)
- id:sz_dr (2018). Calculating the mean and variance of the ratio of random
variables using the Delta method [in Japanese]. *If you are human, think more
now.* https://www.szdrblog.info/entry/2018/11/18/154952
