---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  out.width = 600
)
```

# deltatest

<!-- badges: start -->
<!-- badges: end -->

## Installation

You can install the development version of deltatest from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("hoxo-m/deltatest")
```

## Example

```{r example}
library(dplyr)
library(deltatest)

n_user <- 2000

set.seed(314)
df <- generate_dummy_data(n_user) |> 
  mutate(group = if_else(group == 0L, "control", "treatment")) |>
  group_by(user_id, group) |> 
  summarise(click = sum(metric), pageview = n(), .groups = "drop")
df

deltatest(df, click / pageview, by = group)
```

```{r}
library(ggplot2)

set.seed(314)

p_values <- NULL
for (i in 1:5000) {
  df <- generate_dummy_data(n_user) |> 
    group_by(group) |> 
    summarise(click = sum(metric), pageview = n(), .groups = "drop")
  
  result <- prop.test(df$click, df$pageview, correct = FALSE)
  
  p_values[i] <- result$p.value
}

df <- data.frame(p_value = p_values) |>
  mutate(range = cut(p_value, breaks = seq(0, 1, by = 0.05))) |>
  group_by(range) |>
  summarise(p = factor(ceiling(max(p_value) * 20) / 20), n = n()) |>
  mutate(prop = n / sum(n))
ggplot(df, aes(p, prop)) +
  geom_col() +
  geom_hline(yintercept = 0.05, color = "red") +
  scale_y_continuous(breaks = seq(0, 1, by = 0.02), minor_breaks = NULL) +
  xlab("p-value") + ylab("proportion")
```

```{r}
set.seed(314)

p_values <- NULL
for (i in 1:5000) {
  df <- generate_dummy_data(n_user) |> 
    group_by(user_id, group) |> 
    summarise(click = sum(metric), pageview = n(), .groups = "drop")
  
  result <- deltatest(df, click / pageview, by = group)
  
  p_values[i] <- result$p.value
}

df <- data.frame(p_value = p_values) |>
  mutate(range = cut(p_value, breaks = seq(0, 1, by = 0.05))) |>
  group_by(range) |>
  summarise(p = factor(ceiling(max(p_value) * 20) / 20), n = n()) |>
  mutate(prop = n / sum(n))
ggplot(df, aes(p, prop)) +
  geom_col() +
  geom_hline(yintercept = 0.05, color = "red") +
  scale_y_continuous(breaks = seq(0, 1, by = 0.01), minor_breaks = NULL) +
  xlab("p-value") + ylab("proportion")
```

## References

- Deng, A., Knoblich, U., & Lu, J. (2018). Applying the Delta Method in Metric
Analytics: A Practical Guide with Novel Ideas. *Proceedings of the 24th ACM
SIGKDD International Conference on Knowledge Discovery & Data Mining.*
- Deng, A., Lu, J., & Litz, J. (2017). Trustworthy Analysis of Online A/B Tests:
Pitfalls, challenges and solutions. *Proceedings of the Tenth ACM International
Conference on Web Search and Data Mining.*
- id:sz_dr (2018). Calculating the mean and variance of the ratio of random
variables using the Delta method [in Japanese]. *If you are human, think more
now.* https://www.szdrblog.info/entry/2018/11/18/154952
